---
title: "Machine Learning Project"
author: "M. Chitnis"
date: "July 25, 2015"
output: html_document
---
#Exective Summary: 
####Project goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.The goal is to use the training data to predict the quality of their performance in the test data.

#Get and Clean Data. The original data is from http://groupware.les.inf.puc-rio.br/har
####Dowload the data if not available and review the dataset
```{r echo=FALSE,results='hide'}
#Loading required packages
library(caret)
library(randomForest)
library(e1071)

## Set Working Directory
setwd("/Users/milindchitnis/Documents/My Docs/Data Analytics/RDir/mlearning")

## Get and load data
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

## Check if file exists, else download the data
if(!file.exists("./Data/pml-training.csv")){
  download.file(url = trainUrl, destfile = "./Data/pml-training.csv", method="curl")
}
if(!file.exists("./ProjData/pml-testing.csv")){
  download.file(url = testUrl, destfile = "./Data/pml-testing.csv", method="curl")
}

origTraining <- read.csv("./Data/pml-training.csv")
origTesting <- read.csv("./Data/pml-testing.csv")

## Review Data
names(origTraining)
str(origTraining)
```

####Remove unwanted columns, Zero Variance Data and columns with more than 25% NA
```{r results='hide'}
## Renmove unwanted columns
uIndex <- grep("X|timestamp|user_name|new_window", names(origTraining));
origTraining <- origTraining[,-uIndex];
origTesting <- origTesting[,-uIndex];

## Remove ZeroVariance data
zIndex <- nearZeroVar(origTraining);
origTraining <- origTraining[,-zIndex];
origTesting <- origTesting[,-zIndex];

## Remove columns with more than 25% NA data
origTraining <- origTraining[,colSums(is.na(origTraining)) < nrow(origTraining)*0.75]
origTesting <- origTesting[,colSums(is.na(origTesting)) < nrow(origTesting)*0.75]
```

#Create Models
####Split the cleaned data into 75% for Training and 25% for Testing.
```{r results='hide'}
## Set seed and divide the training use createDataPartition and use 75% data for training
set.seed(12345)
partTrain = createDataPartition(origTraining$classe, p=0.75, list=FALSE)
trainData = origTraining[ partTrain,]
testData =  origTraining[-partTrain,]
```

####Use Random Forest on training data to build prediction model for test data
```{r results='hide'}
## Use randomForest to fit a model for the split trainData 
modelFit <- randomForest(classe ~., data=trainData, p=0.60, list = FALSE)

## Fit the model for the split testData for predicting
modelPred <- predict(modelFit,testData)
```

#Conslusion
####The result of confusionMatrix on the predictive model, has a accuracy of 95% indicating a good fit of the model that was generated.
``` {r}
confusionMatrix(modelPred, testData$classe);
```
####Finally use the model on the set of 20 Test Values that were provided
```{r results='hide'}
## Predict values for original test values of 20 cases
finalPred <- predict(modelFit,origTesting);
```

#Final Submission
####Create output files for the 20 test cases
```{r results='hide'}
## Create output file for submission
answers <- as.vector(finalPred)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("./Output/out_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

## Run functon to create the output
pml_write_files(answers)
```


